---
title: "more_seurat"
author: "alva annett"
date: "1/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(ggplot2)
library(dplyr)
```

# Read Data 

Read10X: 
  data.dir = directory contaning matrix.mtx, genes.tsv (or features.tsv), and barcodes.tsv (from cellranger)
  
```{r read data}

data.pbmc = Read10X(data.dir = "C:\\Users\\alva\\Desktop\\exjobb_master\\data\\filtered_feature_bc_matrix")

```

# Create Seurat object 

CreateSeuratObject: 
  counts = unnormalized raw counts 
  project = project name 
  assay = name of original assay 
  min.cells = min number of cells where a feature is detected 
  min.features = cells with at leas this many features 
  
Filters based on min.cells and min.features. How to choose?? 
 
```{r create seurat object}

pbmc = CreateSeuratObject(counts = data.pbmc, project = "seurat_test", 
                          assay = "RNA", min.cells = 3, min.features = 200)

```

nCount_RNA = total number of molecules 
nFeature = number of unique genes 

```{r meta data}

head(pbmc@meta.data, 5)

``` 

Calculate the percentage of reads comming from mtDNA (indication of dying cell)

```{r add metadata}

pbmc[['percent.mt']] = PercentageFeatureSet(object = pbmc, pattern = "^MT-")

```

```{r QC plot}

# Visualize QC metrics as a violin plot
VlnPlot(pbmc, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, cols ="orange")

```

```{r QC filter}

#subset based on QC metrics 
pbmc <- subset(pbmc, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 20)

```

# Normalize Data 

How to choose scaling factor?? 

```{r norm}

pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

``` 


# Identify Variable featuers 

nfeatures = number of top variable features to return 

Variable features stored in pbmc@assays$RNA@var.features or pbmc[['RNA']]@var.features 

```{r var}

pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
top10 = head(pbmc@assays$RNA@var.features, 10)

# plot variable features with and without labels
plot1 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1


```  

# Scale Data 

Scales all genes in matrix. If featurse argument is omittet only the genes defined by FindVariableFeatures are scaled. 

Scaled data stored in pbmc@assays$RNA@scale.data or pbmc[['RNA']]@scale.data 

```{r scale}

all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)

``` 

# Linear Dimensional Reduction 

Perform on variable features. PCA can be found in pbmc@reductions$pca 

```{r PCA}

pbmc <- RunPCA(pbmc, features = pbmc[['RNA']]@var.features)

#plot PC1 vs PC2 
plot_pca = DimPlot(pbmc, reduction = "pca", cols = "orange")

#plot loadings on first 2 PC
plot_loading = VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")

plot_pca + plot_loading
```

# Determine dimensionality 

Can be done by JackStraw (slow on large datasets) or elbowplot. In this case 14 PC are proceeded with. 

```{r elbow plot}

pbmc <- JackStraw(pbmc, num.replicate = 100)
pbmc <- ScoreJackStraw(pbmc, dims = 1:20)

plot_jack = JackStrawPlot(pbmc, dims = 1:20)
plot_elbow = ElbowPlot(pbmc)

plot_jack + plot_elbow

```

# Clustering 

dims = the dimensions form the PCA 

clusters can be found in pbmc@meta.data or Idents(pbmc)

```{r cluster}

pbmc <- FindNeighbors(pbmc, dims = 1:14)
pbmc <- FindClusters(pbmc, resolution = 0.5)

``` 

# Non-linear dimnesional reduction 

```{r UMAP}
pbmc <- RunUMAP(pbmc, dims = 1:14)

plot_umap = DimPlot(pbmc, reduction = "umap")

```

# Find Markers 

```{r markers}

pbmc.markers <- FindAllMarkers(pbmc, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

#Save top marker for each cluster 
top_marker = pbmc.markers %>% group_by(cluster) %>% top_n(n = 1, wt = avg_logFC)  

FeaturePlot(pbmc, features = top_marker$gene)

```

```{r heatmap}

top5 <- pbmc.markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_logFC)
DoHeatmap(pbmc, features = top5$gene) + NoLegend()
```

